---
title: ""
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---

```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(cache = TRUE)
```

```{r library}
library(tidyverse)
library(here)
library(dplyr)
```

```{r data_read}
NBA = read.csv(here("NBA_player_of_the_week.csv"))

```

1. For each variable identify its type. Which variables are NOT numeric? Does this make sense?

```{r var type}
str(NBA)

# all variables except real values are not numeric. 
# this makes sense since there are no variables with decimal points.
  # 1. cases like weight should be in integer for,
  # 2. cases such as height should be numeric

```

2. Notice that the last 100 or so rows of the dataset have height values in centimeters. The other height values have a specific format, “feet-inches”. Create a new variable called “heightcm” that contains the height of each player in centimeters. This variable should be numeric.

```{r Height_cm}
NBA$Height <- gsub('-', '.', NBA$Height)

vec <- NBA$Height
for (i in 1:length(vec)) {
  if(str_detect(vec[i], 'cm') == TRUE) {
    vec[i] <- str_extract(vec[i],".*(?=cm)")
    NBA$Height_cm[i] <- vec[i]
  
  }
  else {
    NBA$Height_cm[i] <- as.numeric(vec[i])*30.48
  }
}

NBA$Height_cm <- as.numeric(NBA$Height_cm)
```

3. Notice that the last 100 or so rows of the dataset have weight values in kilograms. The rest of the values are in lbs. Create a new variable called “weightkg” that contains the weight of each player in kilograms. This variable should be numeric.

```{r Weight_kg}
vec <- NBA$Weight
for (i in 1:length(vec)) {
  if(str_detect(vec[i], 'kg') == TRUE) {
    vec[i] <- str_extract(vec[i],".*(?=kg)")
    NBA$Weight_kg[i] <- vec[i]
  
  }
  else {
    NBA$Weight_kg[i] <- as.numeric(vec[i])/2.205
  }
}
NBA$Weight_kg <- as.numeric(NBA$Weight_kg)
```

4. Create a new variable called “timesWon” that contains the number of times each Player/Team combination won. Your resulting dataset should only have one row for each Player/Team combination. Note: we will lost the date-related variables with this, but it’s fine!

```{r Times_Won}
NBA1 <- NBA %>% 
  group_by(Player, Team, Conference) %>% 
  summarise(Times_Won = n(), .groups = 'drop')
NBA1
```

5. Which 5 players have won player of the week the most times?

```{r top_5}
NBA1 <- data.frame(NBA1)

NBA1 <-arrange(NBA1,-Times_Won)
head(NBA1)
```
6. Fill in the blanks in the following paragraph after writing code to do the relevant calculations.

This data set contains _____ rows containing information on _____ different positions. Players’ age ranged from _____ to _____, _____% are from the SG position, and the overall average Height value (in centimeters) is _____. The most common team is _____ ( _____%), while the least common team was _____. Although the overall median Weight (in kilograms) was _____, the maximum Weight among East conference players was _____. Overall, _____ percent of these players have won more than 10 times; among the West conference this was _____% and among the East conference this was _____%.

```{r fill}
rows<- nrow(NBA)
rows

positions <- length(unique(NBA$Position))
positions

min_age <- min(NBA$Age)
min_age

max_age <- max(NBA$Age)
max_age

NBA2 <- NBA %>% filter(NBA$Position == "SG")
NBA2 <- length(unique(NBA2$Player))
SG <- (NBA2/(length(unique(NBA$Player))))*100
SG
       
avg_height <- mean(NBA$Height_cm)
avg_height


NBA3 <- NBA %>% 
  group_by(Team) %>% 
  summarise(Most_Common = n()) %>% 
  slice_max(Most_Common, n=1) 
Common <- NBA3$Team
Common

NBA4 <- NBA %>% 
  group_by(Team) %>% 
  summarise(Most_Common = n()) %>% 
  mutate(freq = round(Most_Common/sum(Most_Common)*100,1)) %>% 
  slice_max(Most_Common, n=1)
percent <- NBA4$freq
percent

NBA5 <- NBA %>% 
  group_by(Team) %>% 
  summarise(Most_Common = n()) %>% 
  slice_min(Most_Common, n=1) 
Least <- NBA5$Team
Least

median <- median(NBA$Weight_kg)
median

NBA6 <- NBA %>% 
  filter(NBA$Conference == "East") %>% 
  slice_max(Weight_kg, n=1)
MaxEast <- NBA6$Weight_kg
MaxEast

NBA7 <- NBA1 %>%
  mutate(Times_Won = case_when((NBA1$Times_Won>10) ~ "above10", TRUE ~ "10&below10")) %>%
  group_by(Times_Won) %>%
  summarise(
    individual = n()
  ) %>%
  mutate(percentWin = individual/sum(individual)*100) %>% 
  filter(Times_Won == "above10")


WinPercent <- NBA7$percentWin
WinPercent

NBA8 <- NBA1 %>%
  mutate(Times_Won = case_when((NBA1$Times_Won>10) ~ "above10", TRUE ~ "10&below10")) %>%
  group_by(Times_Won,Conference) %>%
  summarise(
    individual = n(), .groups = 'drop'
  ) %>%
  mutate(percent = individual/sum(individual)*100) %>% 
  filter(Times_Won == 'above10')
NBA8

```

This data set contains 1145 rows containing information on 11 different positions. Players’ age ranged from 19 to 40, 12.41% are from the SG position, and the overall average Height value (in centimeters) is 199.45. The most common team is Los Angeles Lakers ( 6.2%), while the least common team was Nets & Pelicans. Although the overall median Weight (in kilograms) was 102.04, the maximum Weight among East conference players was 147.40. Overall, 5.19 percent of these players have won more than 10 times; among the West conference this was 1.91% and among the East conference this was 1.64%.